Error Reporting Parsers: a Monad Transformer Approach

### 2013-06-24 article.tex
Parsers:  finish up position reporting section and start Parsec comparison section (needs work)

\subsection{}

@@

>>>>

\subsection{Position reporting}
To actually report the position when an error is detected, we introduce a new 
combinator, which I'll call \verb+cut+:
\Matt{a horrible name}
\begin{verbatim}
cut :: String -> Parser Char a -> Parser Char a
cut message parser = 
    getState >>= \p ->
    commit (message, p) parser
\end{verbatim}

It checks the current position, runs a parser, and if the parser fails, reports
an error consiting of a message and the position.

To take advantage of this combinator, we essentially replace every use of 
\verb+commit+ with \verb+cut+:
\begin{verbatim}
application =
    openparen                     >>
    cut eAppOper form             >>= \op ->
    many0 form                    >>= \args ->
    cut eAppClose closeparen      >>
    return (AApp op args)

define =
    check (== "define") symbol    *>
    pure ADefine                 <*>
    cut eDefSym symbol           <*>
    cut eDefForm form

lambda =
    check (== "lambda") symbol     >>
    cut eLamParam opencurly        >>
    many0 symbol                   >>= \params ->
    (if distinct params 
        then return ()
        else cut eLamDupe empty)      >>
    cut eLamPClose closecurly         >>
    cut eLamBody (many1 form)         >>= \bodies ->
    return (ALambda params bodies)
  where
    distinct names = length names == length (nub names)

special = 
    opencurly  *>  cut eSpecial (define <|> lambda)  <*  cut eSpecClose closecurly

form = fmap ASymbol symbol <|> application <|> special

woof = junk *> many0 form <* cut eWoof end
\end{verbatim}

Now the parsers report useful errors:
\begin{verbatim}
*Main> runParser woof " abc 123 "
Left ("woof: unparsed input",(1,6))

*Main> runParser woof " (((abc ())) "
Left ("application: missing operator",(1,10))

*Main> runParser woof "{define const \n  {lambda {x x} \n   x}"
Left ("lambda: duplicate parameter names",(2,15))
\end{verbatim}

@@

\Section{Parsec comparison}
A quick word about Parsec \cite{leijen}, an important, battle-tested parser 
combinator library for Haskell.

One major difference between the approach I've described and Parsec's approach
is with respect to backtracking and error-reporting.  Parsec's default is 
predictive parsing using a single token of lookahead for efficiency.  This can
be changed using its \verb+try+ combinator.

On the other hand, my approach by default uses unlimited lookahead, which can 
be modified using the \verb+throwError+ and \verb+commit+ combinators.
